---
title: "Mental Health in Tech - Analysis Report"
author: "Ankita Pichuka"
output: html_document
---

#Introduction

This project explores mental health trends among tech workers based on the OSMI Mental Health in Tech Survey dataset. The goal is to identify patterns, key factors influencing treatment-seeking behavior, and to build a predictive model.


```{r}
#Lets load the libraries

library(tidyverse)
library(janitor)
library(skimr)
library(ggplot2)

# Load the data
mental_data <- read_csv("mental-health-tech-survey.csv")
```

##Data Cleaning
```{r}
# Clean column names
mental_data <- mental_data %>% clean_names()

# Quick look at structure
glimpse(mental_data)

# Basic summary
skim(mental_data)

# Check missing values
colSums(is.na(mental_data))

# Filter age to a valid range (18 to 65)
mental_data <- mental_data %>% filter(age >= 18 & age <= 65)

# View unique values in key columns
unique(mental_data$gender)
unique(mental_data$treatment)

# Standardize gender values
mental_data <- mental_data %>%
  mutate(gender = str_to_lower(gender),
         gender = case_when(
           str_detect(gender, "female|f") ~ "female",
           str_detect(gender, "male|m") ~ "male",
           TRUE ~ "other"
         ))

# Check cleaned values
table(mental_data$gender)
```

#-----Exploratory Analysis Plots-----
```{r}
#Treatment by Gender
ggplot(mental_data, aes(x = gender, fill = treatment)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Mental Health Treatment by Gender",
       y = "Percentage", x = "Gender") 

# Treatment by Remote Work (Proportional Bar Chart)
ggplot(mental_data, aes(x = remote_work, fill = treatment)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Treatment Seeking by Remote Work Status",
       x = "Remote Work", y = "Proportion") 

# Family History vs Treatment
ggplot(mental_data, aes(x = family_history, fill = treatment)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Effect of Family History on Treatment Seeking",
       x = "Family History", y = "Proportion") +
  theme_minimal()

# Age Distribution
ggplot(mental_data, aes(x = age)) +
  geom_histogram(binwidth = 2, fill = "#4e79a7", color = "white") +
  labs(title = "Age Distribution of Survey Participants",
       x = "Age", y = "Count") +
  theme_minimal()
```

#-----Chi_Square Tests-----
```{r}
#Gender vs Treatment
chisq_gender <- chisq.test(table(mental_data$gender, mental_data$treatment))
chisq_gender

#Remote Work vs Treatment
chisq_remote <- chisq.test(table(mental_data$remote_work, mental_data$treatment))
chisq_remote

#Family history vs Treatment
chisq_family <- chisq.test(table(mental_data$family_history, mental_data$treatment))
chisq_family
```

#----Logistic Regression to Predict Treatment----
```{r}
#Convert treatment to a binary variable
mental_data <- mental_data %>%
  mutate(treatment_bin = ifelse(treatment == "Yes", 1, 0))

#Fit the logistic regression model
log_model <- glm(treatment_bin ~ gender + age + family_history + remote_work,
                 data = mental_data, family = "binomial")

#View model summary
summary(log_model)

#Convert coefficients to odds ratios
exp(coef(log_model))

#Get confidence intervals for those odds ratios
exp(confint(log_model))
```

#Conclusion

- Family history is the strongest predictor of seeking treatment.

- Gender also has a significant association with treatment-seeking.

- Remote work status was not significantly related to treatment.

- Our logistic model gives insights into how demographics and work conditions affect mental health help-seeking behavior.
